---
title: "Mono Lake, Lee Vining, CA: Lake Levels"
author: "Shelby Smith"
date: "12/5/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
library(here)
library(tidyverse)
library(janitor)
library(RColorBrewer)
```

```{r read in data, include=FALSE}
mono_lake<- read_csv(here("data", "mono_lake.csv")) %>% 
  clean_names()
```

```{r, include=FALSE}
#avg vertical change
vert<-mono_lake %>% 
  na.omit(NA) %>% 
  pull(vertical_change)
vert_avg<- mean(vert)
vert_avg
```


```{r, include=FALSE}
mono_lake_viz<- mono_lake %>% 
  ggplot(aes(y=lake_level_feet_above_sea_level, x=year)) +
  geom_line(color="turquoise4")+
#  scale_x_continuous(breaks=seq(from=1850, to=2017, by=10)) could use this to change x axis, but it messes with the annotations at this pont
  
  annotate(geom="point", x=1941, y=6418, size=6, shape=21, fill="transparent")+ #Diversion circle
    annotate(geom="text", x=1953, y=6421, label="1941: Water Diversions to LA Begin", size=2.5)+ #Diversion label
  
  annotate(geom = "point", x=1994, y=6374, size=6, shape=21, fill="transparent")+ #Decision point
    annotate(geom = "text", x=1984, y=6371, label="1994: California State Water Resources Control Board Decision", size=2.5)+ #decision label
  
  geom_hline(yintercept=6360, color="orange", size=.5, linetype='dashed') + #brine shrimp line
    annotate(geom = "text", x=1850, y=6362, 
             label="Decreased availability of brine shrimp as food source for birds", size=2.5, hjust=0)+
  
  geom_hline(yintercept=6392, color="Grey3", size=.5, linetype='dashed') + #stable level line
    annotate(geom="text", x=1850, y=6394, 
             label="Target Lake Level for healthy ecosystem function", size=2.5, hjust=0)+
  
  geom_hline(yintercept=6377, color="red", size=.5, linetype='dashed') + #land bridges line
    annotate(geom = "text", x=1850, y= 6379, 
             label="Land bridges connect mainland predators to island nesting sites", size=2.5, hjust=0 )+
  
  theme_linedraw()+
  labs(title = "Mono Lake Water Levels: 1850-2017",
       x="Year",
       y="Lake Level (feet above sea level)")
```

```{r, echo=FALSE, fig.width = 10}
mono_lake_viz
```

Figure 1: Mono Lake water levels from 1850 to 2017. The blue line shows the lake level measured in feet above sea level, with two specific points highlighted to reflect significant events in the lake's history. Horizontal dashed lines represent ecologically relevant lake levels. Water levels in the lake control salinity: high lake levels dilute the salt content, lowering salinity. High salinity, occurring at low water, decreases food availability for migrating and nesting birds. Data Source: Mono Lake Committee



