---
title: "An Exploration of Political and Environmental Self-Identification"
author: "Shelby Smith"
date: "12/3/2020"
output: html_document
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(here)
library(janitor)
library(kableExtra)
library(broom)

conservation_survey_full<-read.csv(here("data","conservation_survey.csv"))
```

```{r, include=FALSE}
# Create a simplified version of the survey data to only keep variables you need for this task, 'ENV_ID' and 'POL_OR'.
conservation_survey<- conservation_survey_full %>% 
  select("ENV_ID", "POL_OR")

#add a new column that bins the political party into a binary response of either “conservative” (if POL_OR is 1 - 3) or “liberal” (if POL_OR 5 - 7)
political_affiliation<-conservation_survey %>% 
  mutate(affiliation=case_when(
    POL_OR %in% c(1:3) ~ "conservative",
    POL_OR %in% c(5:7) ~ "liberal")) %>% 
  na.omit(political_affiliation, rows="NA") #remove respondents with 4,8,9 political affiliations
```

```{r, include=FALSE}
# Create a finalized table that shows the counts and proportions of “liberal” and “conservative” survey respondents who responded “Yes”, “No” and “Unsure” to the statement “I consider myself an environmentalist.”
pol_aff_counts<-political_affiliation %>% 
  tabyl(affiliation, ENV_ID) 

affiliation_proportions<- pol_aff_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns() %>% 
  rename('Yes'='1',
         'No' = '2',
         'Unsure'='3',
         ' '='affiliation')
```

### Introduction

In this report, we explore how political affiliation influences if a person considers themself an environmentalist or not. The data used in this report is from, *"Influences on charitable giving for conservation"*, which was an online survey administered to over 1000 participants across the United States in August 2017 (Andrews Forest LTER Site and C. Batavia. 2019). We will then use a chi.squared test to determine if there is an association between these two variables. 

### Survey Response Exploration 

Table 1: Survey responses to the questions:  "I consider myself an environmentalist," with outcomes 1 = Yes, 2 = No, 3 = Unsure, and "How do you self-identify politically?" with response values 1 - 3 as “conservative” (strongly to slightly), and values 5 - 7 as “liberal” (slightly to strongly). Shown in parentheses are the actual counts of respondents in these categories. Data source: Andrews Forest LTER Site and C. Batavia. 2019. 

```{r, echo=FALSE}
# show a table in knitted document
final_pol_table<- kbl(affiliation_proportions) %>% 
  kable_paper("hover", full_width=F)
final_pol_table
```

```{r, include=FALSE}

# Make contingency table and run chi-square
pol_aff_ct<- pol_aff_counts %>% 
  column_to_rownames(var="affiliation")
```

```{r, include=FALSE}
# run chi square and make a tidy table
pol_aff_x2<- chisq.test(pol_aff_ct) 
pol_aff_x2
pol_aff_tidy<-tidy(pol_aff_x2)
```

### Hypothesis Test Results and Takeaways

Using a chi.squared test, we conclude that political affiliation has a significant influence on if a person considers themself an environmentalist or not. ($\chi$^2^(`r pol_aff_tidy$parameter`) = `r round(pol_aff_tidy$statistic,2)`, *p* = `r round(pol_aff_tidy$p.value, 3)`). With a 0.05 level of significance, we were able to reject the null hypothesis that political affiliation and environmentalism are independent. 

### Data citation: 
Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c
